Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by \textsc{DigitalSnow} ANR-11-BS02-009, \textsc{KIDICO} ANR-2010-BLAN-0205 and \textsc{PRIMES Labex} ANR-11-LABX-0063/ ANR-11-IDEX-0007 research grants.}

### 2016-01-24 latex/article.tex
minor fix + adding macro for blue text => easier to remove after

Volumetric objects are being more and more popular in many
applications ranging from object modeling and rendering in Computer
Graphics to geometry processing in Medical Imaging or Material
Sciences. When considering large volumetric data, interactive
%visualization of such images (or isovalue surfaces) is a complex
visualization of those objects (or isosurfaces) is a complex
problem. Such issues become even more difficult when dynamic
%volumetric datasets are considered. Beside such isosurface
volumetric datasets are considered. Beside
visualization, we are also interested in performing geometry
processing on the digital object and to explore different parameter
settings of the geometry processing tool.
\color{blue}
Here, we focus on curvature tensor estimation (mean/Gaussian curvature, principal curvatures
directions, \ldots). Most curvature estimators require a parameter fixing the scale
at which the computation is performed. For short, such parameter
(integration radius, convolution kernel size, \ldots) specifies a scale for analyzing the object surface, 
and is naturally related to the amount of perturbation on input
data\todo{DC: ``related to a expected signal/noise ratio?''}. As a consequence, when using
such estimators, exploring different values of this parameter is mandatory.
However, this is usually an offline process, due to the amount of computations that
need to be done.
\normalcolor
%In this work, we present an interactive visualization of volumetric data isosurfaces combined with
%a curvature tensor (mean/Gaussian curvature, principal curvatures
%directions, \ldots) estimation at each point of the digital
%surface.
%Curvature maps on a given isosurface are key information when
%exploring 3D structures in Material Sciences \todo{REF}.

@@

\vspace{0.2cm} \color{blue} In this paper, we first present the
previous works on curvature tensor estimation. Then, we propose na
efficient GPU approach to extract a triangulated isosurface from a
digital object. Finally, we propose a fully-parallel  GPU pipeline to compute
curvature tensor in real time and present our results.
\normalcolor

@@

In \cite{DGCI2013}, estimators are defined based on the 2D digital curvature estimator
$\CurvH{R}$ and the 3D digital mean curvature estimator $\MeanCurvH{R}$ based on
the digital volumetric integral $\VolC(Y,h) \EqDef h^d \MCard(Y)$ (we denote such
quantity $\AreaC(Y,h)$ in dimension 2):
%% Redite
%% \Jaco{Given some point $\vx \in \R^d$, the set $Y$ is the digital intersection of $\DigShape$ with the ball of radius $R/h$ and center $\vx/h$.}
%% where $Y$ is the intersection of $\DigShape$ with a ball
%% $\Ball{R/h}{\vx/h}$ with $\MCard(Y)$, the number of digital points of
%% the intersection between a ball $\Ball{R/h}{\vx/h}$ of digital radius
%% $R/h$ and $\DigShape$ (we denote such quantity $\AreaC(Y,h)$ in
%% dimension 2):
%
\begin{Definition}
  %
  Given the Gauss digitization $Z\EqDef\DSh$ of a shape $\Shape \subset \R^2$
  (or $\R^3$ for the 3D mean curvature estimator), {\em digital curvature
  estimators} are defined for any point $\vx \in \R^2$ (or $\R^3$) as:
  %
  \begin{align}
    \forall 0 < h < R,\quad
    &\CurvH{R}(Z,\vx,h) \EqDef \frac{3\pi}{2R}
    - \frac{3\AreaC(\Ball{R/h}{\vx/h} \cap
    {Z}, h)}{R^3}\,,\label{eq:dig-curvature-estimator-k}\\
    &\MeanCurvH{R}(Z,\vx,h) \EqDef \frac{8}{3R}
    - \frac{4\VolC(\Ball{R/h}{\vx/h} \cap
    Z, h)}{\pi R^4}\,.\label{eq:dig-curvature-estimator-H}
  \end{align}
  %
  where $\Ball{R/h}{\vx/h}$ is the ball of digital radius $R/h$ centered on
  digital point $(\vx/h) \in Z$.
  %
  % \AreaC(\Ball{R/h}{\vx/h} \cap \Body{\DSh}{h}, h) \EqDef h^2 \MCard(
  % \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ and $\VolC(\Ball{R/h}{\vx/h} \cap
  % \Body{\DSh}{h}, h) \EqDef h^3 \MCard( \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ are
  % the number of digital points in the intersection between $\Body{\DSh}{h}$ and
  % the ball $\Ball{R/h}{\vx/h}$ of radius $R/h$ centered on $\vx/h$, scaled by
  % $h^d$.\todo[inline]{J: Hum... pas top...}
  %
  % Similarly, $\hat{\lambda}_1$
  % and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance
  % matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}
%
Estimators of both 2D curvature and 3D mean curvature on digital
objects with volumetric integral information are proved to be multigrid convergent
property \cite{DGCI2013}. For short, when the digital object becomes finer and
finer, \ie when the digitization step $h$ tends to zero, the estimated quantity
on $\Bd{\Body{\DSh}{h}}$ converges (theoretically and experimentally) to the
\color{blue}
associated one on $\Bd{\Shape}$ in $O\left(h^{\frac{1}{3}}\right)$ for
convex shapes with at least $C^3$-boundary and bounded curvature
(setting $R\EqDef kh^\frac{1}{3}$ for some $k\in\R$).
\normalcolor
%
%Since our digital estimators are defined on
%digital shapes, in order to compare results with the exact curvature on
%Euclidean shape, we need to match points on boundary of the digital
%shape and points on the border of the Euclidean one.
%
%Setting $R = kh^\frac{1}{3}$ on convex shapes
%with at least $C^3$-boundary and bounded curvature, where $k$ is a constant
%related to the maximal curvature of the shape, they have:
%%
%\begin{align}
%\forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
%\forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
%&\CurvH{R}(\DSh, \vxH, h) = \Curv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,,\\
%&\MeanCurvH{R}(\DSh, \vxH, h) = \MeanCurv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,.
%\end{align}

In \cite{CVIU2014}, authors have also defined 3D digital principal curvature
estimators $\PrincCurvH{1}{R}$ and $\PrincCurvH{2}{R}$ on $Z\subset\Z^3$ based
on digital moments:
%
\begin{Definition}
  %
  Given the Gauss digitization $Z \EqDef \DSh$ of a shape $\Shape \subset \R^3$,
  {\em 3D digital principal curvature estimators} are defined for any point $\vx
  \in \R^3$ as:
  %
  \begin{align}
    \forall 0 < h < R,\quad
    \PrincCurvH{1}{R}(Z,\vx,h) &\EqDef \frac{6}{\pi {R}^6}(\hat{\lambda}_2 - 3\hat{\lambda}_1) + \frac{8}{5 R} \,,\label{eq:dig-curvature-estimator-k1}\\
    \PrincCurvH{2}{R}(Z,\vx,h) &\EqDef \frac{6}{\pi {R}^6}(\hat{\lambda}_1 - 3\hat{\lambda}_2) + \frac{8}{5 R} \,,
    \label{eq:dig-curvature-estimator-k2}
  \end{align}
  %
  where $\hat{\lambda}_1$ and $\hat{\lambda}_2$ are the two greatest eigenvalues
  of the covariance matrix of the intersection between the digital shape
  $\DigShape$ and a ball of digital radius $R/h$ centered on digital point
  $\vx/h$.
  %
\end{Definition}
%
The covariance matrix needs to compute digital moments of order $0$, order $1$
and order $2$ (see Equation~20 of \cite{CVIU2014} for more details).
% ($\DMom{000}{h}$), of order $1$ ($\DMom{100}{h}$, $\DMom{010}{h}$ and
% $\DMom{001}{h}$) and of order $2$ ($\DMom{200}{h}$, $\DMom{020}{h}$,
% $\DMom{002}{h}$, $\DMom{110}{h}$, $\DMom{011}{h}$ and $\DMom{101}{h}$, see
% Equation~20 of \cite{CVIU2014} for more details).
\color{blue}
These estimators are proven convergent in $O\left(h^{\frac{1}{3}}\right)$ when setting the ball
radius $h$ in $R = kh^\frac{1}{3}$, where $k$ is a constant related to the
maximal curvature of the shape, on convex shapes with at least $C^3$-boundary
and bounded curvature \cite{CVIU2014}.
\normalcolor 
%More formally, it leads to:
%%
%\begin{align}
%  \forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
%  \forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
%  &\PrincCurvH{1}{R}(\DigF{\Shape}{h},\vxH, h) = \PrincCurv{1}(\Shape,\vx) + O\left(h^{\frac{1}{3}}\right) \,,\\
%  &\PrincCurvH{2}{R}(\DigF{\Shape}{h},\vxH, h) = \PrincCurv{2}(\Shape,\vx) + O\left(h^{\frac{1}{3}}\right) \,.
%\end{align}

@@

We first present design principles for the GPU implementation and then
present our approaches.
%\subsection{Design principles}
The general idea is to perform an Integral Invariant computation on
GPU at each vertex of the generated triangulated mesh.
\color{blue}
Since GPU have massively parallel architectures, we can do all those computations in parallel 
to obtain a very efficient implementation. Please note that
\normalcolor
when the LoD criterion is removed, each MC vertex is exactly centered at
a surfel center. At different depth of the octree, MC vertices still
correspond to surfel centers of subsampled versions of the input
object. Hence, Integral Invariant framework defined in
Section~\ref{sec:curv-tens-estim} is consistent with the triangulated
surface obtained on GPU: triangles are used for visualization purposes
but all computations are performed on the digital object $\DSh$ for
estimators
(\ref{eq:dig-curvature-estimator-k}),
(\ref{eq:dig-curvature-estimator-H}), (\ref{eq:dig-curvature-estimator-k1})
and (\ref{eq:dig-curvature-estimator-k2}).

@@

\paragraph{Naive Approach.}
\txtblue{A first simple solution  consists in scanning all digital
points, at a given resolution, lying inside the integration domain}
(Fig.~\ref{fig:approx}-$left$) and then \txtblue{estimating} the
geometrical moment as the sum of the geometrical moments of each elementary
cubes lying in the intersection (see
Eq.~(\ref{eq:dig-curvature-estimator-k}) to (\ref{eq:dig-curvature-estimator-k2})). 
\color{blue}
This is exactly similar to what is done on the CPU. On  the GPU, we can
exploit \emph{mipmap} textures to obtain  multi-resolution information.
\normalcolor
If the input binary object is stored in a
3D texture, GPU hardware constructs a multi-resolution pyramid (mipmap) such that 8
neighboring voxel intensities at level $l$ are averaged to define the voxel
value at level $l+1$. If the level 0 corresponds to the binary input object, at
\color{blue}
a given level $l$, a texture probe at  a point $(x,y,z)$ returns the fraction of $\DSh$ belonging
\normalcolor
to the cube of center $(x,y,z)$ and edge length $2^l$. As a consequence, we can
approximate the volume of $\Ball{R}{\vx}\cap\Shape$ by considering mipmap values
at a given resolution $l$ (Fig.~\ref{fig:approx}-$right$). In this case, errors only
occurs for cells lying inside $\Shape$ (with density 1) not entirely covered by
$\Ball{R}{\vx}$. Furthermore, the \emph{mipmap} texture can be used to design,
using a single texture probe, a fast inclusion test of a given cell $c$ at level
$l$ into the shapes: we say that $c$ is in $\Shape$ if its density (retrieved
from the texture probe at level l) is greater than $1/2$. The idea here is to
mimic a kind of adaptive Gauss digitization process.

@@

\paragraph{Hierarchical Decomposition.}  
\color{blue}
Using the hierarchical nature of the mipmap texture, we could also
consider hierarchical decompositions of $\Ball{R}{\vx}$.
%(Fig.~\ref{fig:approx}-$middle$) 
%or $\Ball{R}{\vx}\cap\Shape$
%(Fig.~\ref{fig:approx2}-$b$). 
The idea is to decompose the
ball into mipmap cells of different resolution in order to limit
the number of texture access (important bottleneck on GPU hardwares)
and to get better approximated quantities. 
%Note that the scenario in
%Fig.~\ref{fig:approx2}-$b$ is clearly the best approach: the
%volume is exact  (\wrt the regular grid case) since geometrical moments
%are additive quantities, and the number of texture fetches is  optimal
%at this point.
%As discussed above, best
%results in terms of texture fetch can be obtained by the approach
%presented in Fig.~\ref{fig:approx2}-$b$. 
%At each vertex, the
%computation algorithm would consist in a recursive decomposition of
%the domain such that a cell is subdivided if it does not belong to
%$\Ball{R}{\vx}\cap\Shape$.
\txtred{On the GPU, computing a hierarchical decomposition of
  $B_R(\vx)$ at each point $\vx$ is highly inefficient since the
  hardware optimizes the parallelism only when the micro-program
  described in the shader has a predictive execution flow. Hence,
  implementing the recursive algorithm (or its de-recursified version)
  requires a lot of branches (conditional \emph{if} tests) in the
  flow.}  We have thus considered a fast multi-resolution
approximation as illustrated in Fig.~\ref{fig:approx}-$middle$: for a
given radius $R$, we precompute a hierarchical decomposition of
$B_R$. Such decomposition is made of octree cells at different
resolutions. At a given MC vertex $\vx$, the algorithm becomes simple
since we just scan the ball cells and estimate the area (or other
moments) from the mipmap values associated to this cell. Note that
these \txtred{precomputed} cells in the $B_R$ decomposition may not be aligned
with mipmap cells. However, we can use GPU %mipmap functionalities
which interpolates mipmap values if we probe at non-discrete
positions. For a given radius $R$, the expected number of cells in
$B_R$ is in $O(\log{R})$. Implementation details on the hierarchical
octree representation can be found in the supplementary material (see
Section \ref{sec:discussion}). \sout{ Still, this method is not fully
exploiting the GPU parallelism.  The hardware optimizes the
parallelism only when the micro-program described in the shader has a
predictive execution flow. Hence, implementing this recursive
algorithm (or its de-recursified version) requires a lot of branches
(conditional \emph{if} tests) in the flow.}\todo{la discussion vient apres}  \normalcolor

>>>>

Volumetric objects are being more and more popular in many
applications ranging from object modeling and rendering in Computer
Graphics to geometry processing in Medical Imaging or Material
Sciences. When considering large volumetric data, interactive
%visualization of such images (or isovalue surfaces) is a complex
visualization of those objects (or isosurfaces) is a complex
problem. Such issues become even more difficult when dynamic
%volumetric datasets are considered. Beside such isosurface
volumetric datasets are considered. Beside
visualization, we are also interested in performing geometry
processing on the digital object and to explore different parameter
settings of the geometry processing tool.
\txtblue{
Here, we focus on curvature tensor estimation (mean/Gaussian curvature, principal curvatures
directions, \ldots). Most curvature estimators require a parameter fixing the scale
at which the computation is performed. For short, such parameter
(integration radius, convolution kernel size, \ldots) specifies a scale for analyzing the object surface,
and is naturally related to the amount of perturbation on input
data\todo{DC: ``related to a expected signal/noise ratio?''}. As a consequence, when using
such estimators, exploring different values of this parameter is mandatory.
However, this is usually an offline process, due to the amount of computations that
need to be done.}
%In this work, we present an interactive visualization of volumetric data isosurfaces combined with
%a curvature tensor (mean/Gaussian curvature, principal curvatures
%directions, \ldots) estimation at each point of the digital
%surface.
%Curvature maps on a given isosurface are key information when
%exploring 3D structures in Material Sciences \todo{REF}.

@@

\vspace{0.2cm} \txtblue{In this paper, we first present the
previous works on curvature tensor estimation. Then, we propose an
efficient GPU approach to extract a triangulated isosurface from a
digital object. Finally, we propose a fully-parallel  GPU pipeline to compute
curvature tensor in real time and present our results.}

@@

In \cite{DGCI2013}, estimators are defined based on the 2D digital curvature estimator
$\CurvH{R}$ and the 3D digital mean curvature estimator $\MeanCurvH{R}$ based on
the digital volumetric integral $\VolC(Y,h) \EqDef h^d \MCard(Y)$ (we denote such
quantity $\AreaC(Y,h)$ in dimension 2):
%% Redite
%% \Jaco{Given some point $\vx \in \R^d$, the set $Y$ is the digital intersection of $\DigShape$ with the ball of radius $R/h$ and center $\vx/h$.}
%% where $Y$ is the intersection of $\DigShape$ with a ball
%% $\Ball{R/h}{\vx/h}$ with $\MCard(Y)$, the number of digital points of
%% the intersection between a ball $\Ball{R/h}{\vx/h}$ of digital radius
%% $R/h$ and $\DigShape$ (we denote such quantity $\AreaC(Y,h)$ in
%% dimension 2):
%
\begin{Definition}
  %
  Given the Gauss digitization $Z\EqDef\DSh$ of a shape $\Shape \subset \R^2$
  (or $\R^3$ for the 3D mean curvature estimator), {\em digital curvature
  estimators} are defined for any point $\vx \in \R^2$ (or $\R^3$) as:
  %
  \begin{align}
    \forall 0 < h < R,\quad
    &\CurvH{R}(Z,\vx,h) \EqDef \frac{3\pi}{2R}
    - \frac{3\AreaC(\Ball{R/h}{\vx/h} \cap
    {Z}, h)}{R^3}\,,\label{eq:dig-curvature-estimator-k}\\
    &\MeanCurvH{R}(Z,\vx,h) \EqDef \frac{8}{3R}
    - \frac{4\VolC(\Ball{R/h}{\vx/h} \cap
    Z, h)}{\pi R^4}\,.\label{eq:dig-curvature-estimator-H}
  \end{align}
  %
  where $\Ball{R/h}{\vx/h}$ is the ball of digital radius $R/h$ centered on
  digital point $(\vx/h) \in Z$.
  %
  % \AreaC(\Ball{R/h}{\vx/h} \cap \Body{\DSh}{h}, h) \EqDef h^2 \MCard(
  % \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ and $\VolC(\Ball{R/h}{\vx/h} \cap
  % \Body{\DSh}{h}, h) \EqDef h^3 \MCard( \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ are
  % the number of digital points in the intersection between $\Body{\DSh}{h}$ and
  % the ball $\Ball{R/h}{\vx/h}$ of radius $R/h$ centered on $\vx/h$, scaled by
  % $h^d$.\todo[inline]{J: Hum... pas top...}
  %
  % Similarly, $\hat{\lambda}_1$
  % and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance
  % matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}
%
Estimators of both 2D curvature and 3D mean curvature on digital
objects with volumetric integral information are proved to be multigrid convergent
property \cite{DGCI2013}. For short, when the digital object becomes finer and
finer, \ie when the digitization step $h$ tends to zero, the estimated quantity
on $\Bd{\Body{\DSh}{h}}$ converges (theoretically and experimentally) to the
\txtblue{
associated one on $\Bd{\Shape}$ in $O\left(h^{\frac{1}{3}}\right)$ for
convex shapes with at least $C^3$-boundary and bounded curvature
(setting $R\EqDef kh^\frac{1}{3}$ for some $k\in\R$).}
%
%Since our digital estimators are defined on
%digital shapes, in order to compare results with the exact curvature on
%Euclidean shape, we need to match points on boundary of the digital
%shape and points on the border of the Euclidean one.
%
%Setting $R = kh^\frac{1}{3}$ on convex shapes
%with at least $C^3$-boundary and bounded curvature, where $k$ is a constant
%related to the maximal curvature of the shape, they have:
%%
%\begin{align}
%\forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
%\forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
%&\CurvH{R}(\DSh, \vxH, h) = \Curv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,,\\
%&\MeanCurvH{R}(\DSh, \vxH, h) = \MeanCurv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,.
%\end{align}

In \cite{CVIU2014}, authors have also defined 3D digital principal curvature
estimators $\PrincCurvH{1}{R}$ and $\PrincCurvH{2}{R}$ on $Z\subset\Z^3$ based
on digital moments:
%
\begin{Definition}
  %
  Given the Gauss digitization $Z \EqDef \DSh$ of a shape $\Shape \subset \R^3$,
  {\em 3D digital principal curvature estimators} are defined for any point $\vx
  \in \R^3$ as:
  %
  \begin{align}
    \forall 0 < h < R,\quad
    \PrincCurvH{1}{R}(Z,\vx,h) &\EqDef \frac{6}{\pi {R}^6}(\hat{\lambda}_2 - 3\hat{\lambda}_1) + \frac{8}{5 R} \,,\label{eq:dig-curvature-estimator-k1}\\
    \PrincCurvH{2}{R}(Z,\vx,h) &\EqDef \frac{6}{\pi {R}^6}(\hat{\lambda}_1 - 3\hat{\lambda}_2) + \frac{8}{5 R} \,,
    \label{eq:dig-curvature-estimator-k2}
  \end{align}
  %
  where $\hat{\lambda}_1$ and $\hat{\lambda}_2$ are the two greatest eigenvalues
  of the covariance matrix of the intersection between the digital shape
  $\DigShape$ and a ball of digital radius $R/h$ centered on digital point
  $\vx/h$.
  %
\end{Definition}
%
The covariance matrix needs to compute digital moments of order $0$, order $1$
and order $2$ (see Equation~20 of \cite{CVIU2014} for more details).
% ($\DMom{000}{h}$), of order $1$ ($\DMom{100}{h}$, $\DMom{010}{h}$ and
% $\DMom{001}{h}$) and of order $2$ ($\DMom{200}{h}$, $\DMom{020}{h}$,
% $\DMom{002}{h}$, $\DMom{110}{h}$, $\DMom{011}{h}$ and $\DMom{101}{h}$, see
% Equation~20 of \cite{CVIU2014} for more details).
\txtblue{
These estimators are proven convergent in $O\left(h^{\frac{1}{3}}\right)$ when setting the ball
radius $h$ in $R = kh^\frac{1}{3}$, where $k$ is a constant related to the
maximal curvature of the shape, on convex shapes with at least $C^3$-boundary
and bounded curvature \cite{CVIU2014}.}
%More formally, it leads to:
%%
%\begin{align}
%  \forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
%  \forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
%  &\PrincCurvH{1}{R}(\DigF{\Shape}{h},\vxH, h) = \PrincCurv{1}(\Shape,\vx) + O\left(h^{\frac{1}{3}}\right) \,,\\
%  &\PrincCurvH{2}{R}(\DigF{\Shape}{h},\vxH, h) = \PrincCurv{2}(\Shape,\vx) + O\left(h^{\frac{1}{3}}\right) \,.
%\end{align}

@@

We first present design principles for the GPU implementation and then
present our approaches.
%\subsection{Design principles}
The general idea is to perform an Integral Invariant computation on
GPU at each vertex of the generated triangulated mesh.
\txtblue{
Since GPU have massively parallel architectures, we can do all those computations in parallel
to obtain a very efficient implementation. Please note that}
when the LoD criterion is removed, each MC vertex is exactly centered at
a surfel center. At different depth of the octree, MC vertices still
correspond to surfel centers of subsampled versions of the input
object. Hence, Integral Invariant framework defined in
Section~\ref{sec:curv-tens-estim} is consistent with the triangulated
surface obtained on GPU: triangles are used for visualization purposes
but all computations are performed on the digital object $\DSh$ for
estimators
(\ref{eq:dig-curvature-estimator-k}),
(\ref{eq:dig-curvature-estimator-H}), (\ref{eq:dig-curvature-estimator-k1})
and (\ref{eq:dig-curvature-estimator-k2}).

@@

\paragraph{Naive Approach.}
\txtblue{A first simple solution  consists in scanning all digital
points, at a given resolution, lying inside the integration domain}
(Fig.~\ref{fig:approx}-$left$) and then \txtblue{estimating} the
geometrical moment as the sum of the geometrical moments of each elementary
cubes lying in the intersection (see
Eq.~(\ref{eq:dig-curvature-estimator-k}) to (\ref{eq:dig-curvature-estimator-k2})).
\txtblue{
This is exactly similar to what is done on the CPU. On  the GPU, we can
exploit \emph{mipmap} textures to obtain  multi-resolution information.}
If the input binary object is stored in a
3D texture, GPU hardware constructs a multi-resolution pyramid (mipmap) such that 8
neighboring voxel intensities at level $l$ are averaged to define the voxel
value at level $l+1$. If the level 0 corresponds to the binary input object, at
\txtblue{
a given level $l$, a texture probe at  a point $(x,y,z)$ returns the fraction of $\DSh$ belonging}
to the cube of center $(x,y,z)$ and edge length $2^l$. As a consequence, we can
approximate the volume of $\Ball{R}{\vx}\cap\Shape$ by considering mipmap values
at a given resolution $l$ (Fig.~\ref{fig:approx}-$right$). In this case, errors only
occurs for cells lying inside $\Shape$ (with density 1) not entirely covered by
$\Ball{R}{\vx}$. Furthermore, the \emph{mipmap} texture can be used to design,
using a single texture probe, a fast inclusion test of a given cell $c$ at level
$l$ into the shapes: we say that $c$ is in $\Shape$ if its density (retrieved
from the texture probe at level l) is greater than $1/2$. The idea here is to
mimic a kind of adaptive Gauss digitization process.

@@

\paragraph{Hierarchical Decomposition.}
\txtblue{
Using the hierarchical nature of the mipmap texture, we could also
consider hierarchical decompositions of $\Ball{R}{\vx}$.
%(Fig.~\ref{fig:approx}-$middle$)
%or $\Ball{R}{\vx}\cap\Shape$
%(Fig.~\ref{fig:approx2}-$b$).
The idea is to decompose the
ball into mipmap cells of different resolution in order to limit
the number of texture access (important bottleneck on GPU hardwares)
and to get better approximated quantities.
%Note that the scenario in
%Fig.~\ref{fig:approx2}-$b$ is clearly the best approach: the
%volume is exact  (\wrt the regular grid case) since geometrical moments
%are additive quantities, and the number of texture fetches is  optimal
%at this point.
%As discussed above, best
%results in terms of texture fetch can be obtained by the approach
%presented in Fig.~\ref{fig:approx2}-$b$.
%At each vertex, the
%computation algorithm would consist in a recursive decomposition of
%the domain such that a cell is subdivided if it does not belong to
%$\Ball{R}{\vx}\cap\Shape$.
\txtred{On the GPU, computing a hierarchical decomposition of
  $B_R(\vx)$ at each point $\vx$ is highly inefficient since the
  hardware optimizes the parallelism only when the micro-program
  described in the shader has a predictive execution flow. Hence,
  implementing the recursive algorithm (or its de-recursified version)
  requires a lot of branches (conditional \emph{if} tests) in the
  flow.}  We have thus considered a fast multi-resolution
approximation as illustrated in Fig.~\ref{fig:approx}-$middle$: for a
given radius $R$, we precompute a hierarchical decomposition of
$B_R$. Such decomposition is made of octree cells at different
resolutions. At a given MC vertex $\vx$, the algorithm becomes simple
since we just scan the ball cells and estimate the area (or other
moments) from the mipmap values associated to this cell. Note that
these \txtred{precomputed} cells in the $B_R$ decomposition may not be aligned
with mipmap cells. However, we can use GPU %mipmap functionalities
which interpolates mipmap values if we probe at non-discrete
positions. For a given radius $R$, the expected number of cells in
$B_R$ is in $O(\log{R})$. Implementation details on the hierarchical
octree representation can be found in the supplementary material (see
Section \ref{sec:discussion}). \sout{ Still, this method is not fully
exploiting the GPU parallelism.  The hardware optimizes the
parallelism only when the micro-program described in the shader has a
predictive execution flow. Hence, implementing this recursive
algorithm (or its de-recursified version) requires a lot of branches
(conditional \emph{if} tests) in the flow.}\todo{la discussion vient apres}}
